---
title: "withLenghScaledTPM"
author: "Daniel Hanhart"
date: "3/22/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(Seurat)
library(liana)

#library(devtools)
#remotes::install_github('saezlab/liana@0.1.7')

```

## Overview



## load Data


```{r, include=FALSE}
if(FALSE){
  bone_metadata <- read.table("./../local/bone_metadata.tsv", sep = "\t", header = TRUE)
  
  ## only from 2D
  #bone_metadata <- bone_metadata %>% dplyr::filter(Condition == "2D")
  bone_counts <- read_delim(file =  "./../local/bone_tximport_lengthScaledTPM.tsv", delim = "\t", col_names = TRUE) #%>% tibble::column_to_rownames("gene")
  
  rapalink_metadata <- read.table("./../local/rapalink_metadata.txt", sep = "\t", header = TRUE)
  rapalink_counts <- read_delim(file =  "./../local/rapalink_tximport_lengthScaledTPM.tsv", delim = "\t", col_names = TRUE)# %>% tibble::column_to_rownames("gene")
  
  
  ## combine count tables, (in that case no NA produced!)
  counts <- bone_counts %>% full_join(rapalink_counts, by = "gene") %>% column_to_rownames("gene")
  
  ## unify metadata
  m_1 <- bone_metadata %>% dplyr::filter(Condition == "2D")%>% dplyr::rename(sample = GS_Code) %>% dplyr::select("sample", "Timecourse", "Type", "Batch" ) %>% mutate(treatment = NA)  %>% mutate(dataset = "bone", Condition = NA)
  m_2 <- rapalink_metadata %>% dplyr::rename(sample = GenomeScanID, treatment = Treatment, Type = Model) %>% mutate(Timecourse = NA, Batch = NA) %>% dplyr::select("sample", "Timecourse", "Type", "Batch", "treatment" , "Condition") %>% mutate(dataset = "rapalink")
  
  metadata <- rbind(m_1, m_2)
  # reorder metadata based on count matrix
  
  metadata <- metadata[match(metadata$sample, colnames(counts)),]
  rownames(metadata) <- metadata$sample

}

```

## create Seurat object

```{r}
#sobj <- CreateSeuratObject(counts = counts, project = "bone_met", assay = "RNA", meta.data = metadata, min.cells = 0, min.features = 0)
#saveRDS(sobj, "./../local/SeuratObject.rds")
```

## visualize data / structure

```{r}
sobj <- readRDS("./../local/SeuratObject.rds")

a <- sobj %>%  FindVariableFeatures() %>% ScaleData() %>% RunPCA() %>% RunUMAP(dims = 1:30)

DimPlot(a, group.by = "dataset")
DimPlot(a, group.by = "Timecourse")
DimPlot(a, group.by = "Type")

## Bone separates by Batch
bone <- subset(sobj, dataset == "bone")
bone <- bone %>%  FindVariableFeatures() %>% ScaleData() %>% RunPCA(npcs = 40) %>% RunUMAP(dims = 1:30)
DimPlot(bone, group.by = "dataset")
DimPlot(bone, group.by = "Timecourse")
DimPlot(bone, group.by = "Batch")


## rapalink separates by TYPE (BM18 vs LAPC9)
rapalink <- subset(sobj, dataset == "rapalink")
rapalink <- rapalink %>%  FindVariableFeatures() %>% ScaleData() %>% RunPCA(npcs = 35) %>% RunUMAP(dims = 1:30)
DimPlot(rapalink, group.by = "Condition")
DimPlot(rapalink, group.by = "treatment")
DimPlot(rapalink, group.by = "Type")


rm(a, bone, rapalink)
```

## Performe cell-cell communication inference

### HB vs LAPC9 intact

```{r}

get_communication <- function(object, type = "HB", time = 0){
  
  selected_samples <- object@meta.data %>% dplyr::filter((Type == type | Type == "LAPC9")) %>%
  dplyr::filter((is.na(treatment) | treatment == "VEH")) %>%
  dplyr::filter((is.na(Timecourse) | Timecourse == time)) %>%
  dplyr::filter((is.na(Condition) | Condition == "Intact"))

  sobj_specific <- object[,colnames(object) %in% selected_samples$sample]
  
  sobj_specific <- SetIdent(sobj_specific, value = sobj_specific@meta.data$Type )
  
  used_tools <- c("connectome", "logfc", "natmi", "sca", "cellphonedb", "cytotalk",
                      "call_cellchat")
  # "call_italk"
  liana_results <- liana_wrap(sobj_specific, method = used_tools, base = NaN)
  
  
  
    TF <-  sapply(liana_results, function(x){all(c("source", "target") %in% names(x))})
    ## aggregate methods. aggregated by a re-implementation of the RRA method from
    ## RobustRankAggreg. Values can be interpreted as p-values
    liana_results_single_score <- liana_results[TF] %>% liana_aggregate()
    
    p.adj <- liana_results_single_score$aggregate_rank * length(names(liana_results))
    p.adj <- p.adjust(p.adj, "BH")
    liana_results_single_score$p.adj <- p.adj

  return(list(single_score = liana_results_single_score, all_scores = liana_results))
}

```



```{r}
if(FALSE){
sobj <- readRDS("./../local/SeuratObject.rds")

res <- list()

res[["0"]] <- get_communication(sobj, type = "HB", time = 0)
res[["7"]] <- get_communication(sobj, type = "HB", time = 7)
res[["14"]] <- get_communication(sobj, type = "HB", time = 14)
res[["21"]] <- get_communication(sobj, type = "HB", time = 21)
#res[["28"]] <- get_communication(sobj, type = "HB", time = 28)

saveRDS(res, "./../local/cc_res_HB_LAPC9.rds")
}

```

### Evaluate HB_LAPC9 res

```{r}
sobj <- readRDS("./../local/SeuratObject.rds")

res <- readRDS("./../local/cc_res_HB_LAPC9.rds")
res[["28"]] <- NULL

res_single_score <- do.call(rbind, lapply(names(res), function(x){
  print(x)
  r <- res[[x]][[1]]

  r <- r %>% mutate(time = x) %>% dplyr::select("source", "target", "ligand", "receptor", "aggregate_rank", "p.adj", "mean_rank", "time")
  r
}))

res_single_score$time <- as.numeric(res_single_score$time)

## remove autocrine signals

res_paracrine <- res_single_score[res_single_score$source != res_single_score$target,]
res_paracrine$signal <- paste0(res_paracrine$ligand,"_", res_paracrine$receptor)



```


We see that the number of predicted paracrine interactions seems to be the highest after 21 days.

```{r}
## Paractine summary

DF <- res_paracrine %>% group_by(time) %>% dplyr::filter(aggregate_rank < 0.05) %>%summarize(n=n())


print("HB_LAPC9 Paracrine signaling summary")
DF
```


When we look at the p_value distribution we do not see a very low number of p-values in the range of 0-0.05, which most likely mean that they are "false positives". We anyway can examine them

```{r}
#### all values

res_single_score$time <- factor(res_single_score$time, levels = sort(unique(res_single_score$time)))

ggplot(res_single_score, aes(aggregate_rank, fill = time)) + 
  geom_histogram(breaks = seq(0,1,0.05), color = "black") +
  theme_bw(base_size = 12) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  facet_wrap(vars(time), ncol = 2) + ggtitle("aggregation scores histogram")

#### only paracrine signaling

res_paracrine$time <- factor(res_paracrine$time, levels = sort(unique(res_single_score$time)))

ggplot(res_paracrine, aes(aggregate_rank, fill = time)) + 
  geom_histogram(breaks = seq(0,1,0.05), color = "black") +
  theme_bw(base_size = 12) +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  facet_wrap(vars(time), ncol = 2) + ggtitle("aggregation scores histogram (paracrine)")


```


### Venn Diagram between timepoints (all LR pairs)

```{r}

res_paracrine$ID <- paste0(res_paracrine$source,"_",res_paracrine$target,"_",res_paracrine$ligand,"_", res_paracrine$receptor)

venn_list <- list() 
for(t in unique(res_paracrine$time)){
  venn_list[[t]] <- res_paracrine %>% filter(aggregate_rank < 0.05) %>% filter(time == t) %>%  arrange(aggregate_rank)  %>% pull(ID)
}

library("ggVennDiagram")
p <- ggVennDiagram(venn_list)
p

res_paracrine_in_all_timepoints <- res_paracrine %>% filter(ID %in%  Reduce(intersect, venn_list))
DT::datatable(res_paracrine_in_all_timepoints)
#Reduce(intersect, venn_list)

```


### aggregate ranks over time points

```{r}
res_single_score$ID <- paste0(res_single_score$source,"_",res_single_score$target,"_",res_single_score$ligand,"_", res_single_score$receptor)
ranked_list <- list()
for(t in unique(res_single_score$time)){
  ranked_list[[t]] <- res_single_score %>% filter(time == t) %>%  arrange(aggregate_rank)  %>% pull(ID)
}

library(RobustRankAggreg)
res_single_score_RRA <- aggregateRanks(ranked_list, method = "RRA") %>% rename(ID = Name) %>% right_join(res_single_score, by = "ID", multiple = "all")


res_single_score_RRA_paracine <- res_single_score_RRA[res_single_score_RRA$source != res_single_score_RRA$target,]

DT::datatable(res_single_score_RRA)


################################################################################
## only secreted, membrane bound

HPA <- read.table("./../local/proteinatlas_8d579e31.tsv", sep = "\t", header = TRUE, quote = "\"")

HPA_sub <- HPA %>% dplyr::select("Gene", "Subcellular.location", "Secretome.location")

## selct interesting secretion locations: (for ligand select if secreted or at least membrane bound)
unique(HPA_sub$Secretome.location)
TF_1 <- grepl( "Secreted", HPA_sub$Secretome.location)
TF_2 <- grepl( "Immunoglobulin genes", HPA_sub$Secretome.location)
TF_3 <- grepl( "Intracellular and membrane", HPA_sub$Secretome.location)
TF <- (TF_1 | TF_2 | TF_3)
secreted_genes <- HPA_sub %>% filter(TF) %>% pull("Gene")
rm(TF,TF_1, TF_2, TF_3)
## for receptor use only the onces located at the plasma membrane
#unique(unlist(strsplit(HPA_sub$Subcellular.location, split = ",")))
## only choose the onces locataed at plasma membrane. The other terms are not applicable
TF <- grepl("Plasma membrane", HPA_sub$Subcellular.location)
table(TF)
plasma_membrane_genes <- HPA_sub %>% filter(TF) %>% pull("Gene")
rm(TF)

## subset results to only contain receptor at plasma membrane and ligand that are secreted or at the plasma membrane
TF <- (res_single_score_RRA_paracine$ligand %in% secreted_genes  & res_single_score_RRA_paracine$receptor %in% plasma_membrane_genes)
table(TF)

res <- res_single_score_RRA_paracine %>% filter(TF)


df <- DT::datatable(res)
df



DF <- res
## select only HB and LAPC9 for expression table
selected_samples <- sobj@meta.data %>% dplyr::filter((Type == "HB" | Type == "LAPC9")) %>%
  dplyr::filter((is.na(treatment) | treatment == "VEH")) %>%
  dplyr::filter((is.na(Condition) | Condition == "Intact"))

sobj_specific <- sobj[,colnames(sobj) %in% selected_samples$sample]
sobj_specific$Type_Timecourse <- paste0(sobj_specific$Type, "_", sobj_specific$Timecourse)

features <- res %>% arrange(Score) %>% mutate(signal = paste0(ligand,"_",receptor)) %>% pull(signal) %>% unique() %>% strsplit("_") %>% unlist(recursive = TRUE)



expr <- AverageExpression(sobj_specific, features = features, group.by = c("Type_Timecourse"), slot = "counts")$RNA %>% as.data.frame() %>%rownames_to_column("gene")

## expression table
DT::datatable(expr)

# ## merge with expression with res
# ligand_expr <- expr
# names(ligand_expr) <- paste0("ligandExpr_", names(ligand_expr))
# names(ligand_expr)[1] <- "ligand" 
# res <- res %>% left_join(ligand_expr, by = "ligand")
# 
# receptor_expr <- expr
# names(receptor_expr) <- paste0("receptorExpr_", names(receptor_expr))
# names(receptor_expr)[1] <- "receptor" 
# res <- res %>% left_join(receptor_expr, by = "receptor")


## create and save top 10 list:

res_top20 <- res %>% dplyr::select("source", "target", "ligand", "receptor", "Score") %>% distinct() %>% head(n = 20)
library(webshot)
library(htmlwidgets)
df <- DT::datatable(res_top20)
saveWidget(df, "./../local/top20_LAPC9_HB_RRA_over_time_secreted_plasma_membrane_only.html")

```

### Venn Diagram between timepoints (all LR pairs)

```{r}

venn_list <- list() 
for(t in unique(res$time)){
  venn_list[[t]] <- res %>% filter(aggregate_rank < 0.05) %>% filter(time == t) %>%  arrange(aggregate_rank)  %>% pull(ID)
}

library("ggVennDiagram")
p <- ggVennDiagram(venn_list) + ggtitle("Paracrine communication (membrane bound, secreted)")
p

ggsave("./../local/venn_diagram_HB_LAPC9_membrane_secreted.pdf")
#res_paracrine_in_all_timepoints <- res_paracrine %>% filter(ID %in%  Reduce(intersect, venn_list))
#DT::datatable(res_paracrine_in_all_timepoints)
#Reduce(intersect, venn_list)

```





### Investigate with only cellchat (in progress)

```{r}
run_cellchat <- function(object, type = "HB", time = 0){
   library(CellChat)
  
  selected_samples <- object@meta.data %>% dplyr::filter((Type == type | Type == "LAPC9")) %>%
  dplyr::filter((is.na(treatment) | treatment == "VEH")) %>%
  dplyr::filter((is.na(Timecourse) | Timecourse == time)) %>%
  dplyr::filter((is.na(Condition) | Condition == "Intact"))

  sobj_specific <- object[,colnames(object) %in% selected_samples$sample]
 
  cc <- createCellChat(sobj_specific)
  cc <- setIdent(cc, ident.use = "Type")
  
  levels(cc@idents)
  as.numeric(table(cc@idents))
  
  CellChatDB <- CellChatDB.human
  cc@DB <- CellChatDB
  cc <- subsetData(cc)
  
  cc <- identifyOverExpressedGenes(cc)
  cc <- identifyOverExpressedInteractions(cc)
  
  cc <- computeCommunProb(cc)
  cc <- filterCommunication(cc, min.cells = 0)
  
  df.net <- subsetCommunication(cc)
  return(df.net)
}




```

```{r}
if(FALSE){
  sobj <- readRDS("./../local/SeuratObject.rds")
  object <- sobj  
  
  res_cellchat <- list()
  
  res_cellchat[["0"]] <- run_cellchat(sobj, type = "HB", time = 0)
  res_cellchat[["7"]] <- run_cellchat(sobj, type = "HB", time = 7)
  res_cellchat[["14"]] <- run_cellchat(sobj, type = "HB", time = 14)
  res_cellchat[["21"]] <- run_cellchat(sobj, type = "HB", time = 21)
  #res[["28"]] <- get_communication(sobj, type = "HB", time = 28)
  
  saveRDS(res_cellchat, "./../local/res_cellchat_HB_LAPC9.rds")
}
```

```{r}
res_cellchat <- readRDS("./../local/res_cellchat_HB_LAPC9.rds")

a <- do.call(rbind, res_cellchat)

res_cellchat_paracrine <- res_cellchat[res_cellchat$source != res_cellchat$target,]

```
